Modelling the Lymphatic Metastatic Progression Pathways of OPSCC from Multi-Institutional Datasets

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

In total, the dataset contains 681 patients newly diagnosed with \gls{opscc} that were treated with either definitive radiotherapy or neck dissection. Pathologically assessed post-operative \gls{lnl} involvement was available for 393 of patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities. If multiple modalities were used to diagnose a patient's lymphatic involvement and conflicting conclusions were drawn from this, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately. To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated below.

\noindent
\begin{center}
    \begin{tabular}{|l|rr|}
        \hline
        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity} \\
        \hline
        \acrshort{ct} & 76\% & 81\% \\
        \acrshort{pet} & 86\% & 79\% \\
        \gls{mri} & 63\% & 81\% \\
        \acrshort{fna} & 98\% & 80\% \\
        Pathology & $\approx$ 100\% & $\approx$ 100\% \\
        \hline
    \end{tabular}
\end{center}

>>>>

In total, the dataset contains \variable{output/num_patients.txt} patients newly diagnosed with \gls{opscc} that were treated with either definitive radiotherapy or neck dissection. Pathologically assessed post-operative \gls{lnl} involvement was available for 393 of patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities. If multiple modalities were used to diagnose a patient's lymphatic involvement and conflicting conclusions were drawn from this, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately. To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.

\begin{table}
\centering
    \begin{tabular}{|l|rr|}
        \hline
        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity} \\
        \hline
        \acrshort{ct} & 76\% & 81\% \\
        \acrshort{pet} & 86\% & 79\% \\
        \gls{mri} & 63\% & 81\% \\
        \acrshort{fna} & 98\% & 80\% \\
        Pathology & $\approx$ 100\% & $\approx$ 100\% \\
        \hline
    \end{tabular}
    \caption{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
    \label{table:sens_spec}
\end{table}
